Automatic Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{Research supported by NSF OAC \#1940233 and \#1940322.} }

### 2021-04-08 paper/paper_new.tex
Edited Sections 6 & 7

\begin{figure}[b]
  \centering
  \includegraphics[width=0.9\linewidth]{images/none1_crop} \\
  \vspace{3mm}
  \includegraphics[width=0.9\linewidth]{images/double1_crop}
  \caption{A fish that was not detected (top) and a fish that was detected twice (bottom).}
  \label{fig:fish_detect_wrong}
\end{figure}

@@

\section{Discussion}
Overall our results show a proof of concept and offer a path forward for using object detection technology, enhanced by image informatics techniques, to improve and enrich metadata that enables researchers to perform advanced specimen image analysis and investigate scientific research questions on a scale that has never been pursued before. Our investigation has thus far focused on fish as the specimen of study. Fish are vertebrate animals (phylum Chordata), with over \(34,000\) known unique species~\cite{fishbase}, and likely many more undiscovered. Species names are merely labels, and the discovery of species variation depends on both genotype and phenotype information. The ability to analyze thousands of images of a single fish species computationally can lead to new discoveries that are impossible to pursue manually. Digital library researchers have been concerned with computationally extracting image features, using content based image retrieval methods, and work by Toress~\cite{torres2006digital}, while over 15 years old, demonstrates the challenges and opportunity to automatically generate useful metadata. Efforts to integrate such automatic metadata generation methods into digital library workflows and architectures still seem limited.
%content based information retrieval (CBIR) for digital library images\cite{ivanova2011content}
%Link above is from the EU counterpart to JCDL, and trying to weave in, just focuses on color semantics, if we want to get another citation in there.
This is likely due to the diversity of image shapes, sizes and the inconsistent configurations of specimens, labels, rulers, etc.\ within them. Object detection as explored in our research, working with an established architecture, is applicable the larger world of biodiversity, well beyond fish, to include other fauna and flora, art and artefacts, and other digitized images made accessible in for scientific and scholarly research. And following object detection, one can apply pixel analysis to computer many more higher order properties from the initial segmentation information.
%can we say something here.

Digital libraries serve to collect, provide access to, and archive rich collections of a wide array of materials. Many digital libraries interconnect with open repository, supporting FAIR (Findable, Accessible, Interoperable, and Reuseable)~\cite{wilkinson2016fair}. In discussing the future of digital library services \cite{fox2021building} underscores the need to prepare for and test ML applications. The work presented here, within the context of the BGNN project, demonstrates a clear need for improved metadata associated with specimen image collections. Much effort, time and money have been put into photographing and digitizing physical specimen, but without detailed and complete metadata properties the utility of these repositories for advanced computation analysis and ML is limited. Since it is prohibitively time intensive to generate all the pertinent properties by hand, automated techniques are needed to generate this missing metadata at scale.

\section{Conclusion}
In this paper we presented such a metadata generation program. Using ML and image informatics tools and techniques, it is able to locate, bitmask and analyze specimen (currently limited to fish) in collection images with a high degree of accuracy. It produces 6 of the 22 core BGNN metadata properties~\cite{leipzig2021biodiversity}, as well as image contrast, bounding boxes and scale and length information. Testing this program on \(7,244\) images from the INHS dataset~\cite{INHS}, we see that the vast majority of the resulting metadata is correct within a tolerance of a few percentage points, and in some cases contains fewer mistakes than the manually generated validation data. Through further refinement and generalizing beyond only INHS images, we aim to create a tool that can be distributed to specimen image collection curators to correct the metadata sparsity that precipitated this work.
\subsection{Future Work}
The most pressing next step is to refine the pixel analysis thresholding process so that the entirety of even light colored fish are marked as foreground in the mask. A deficiency of the current process is that it only operates on single channel intensity. Some of the lightest tails appear yellow in hue to the human eye and easily distinguishable, but when compressed to a single intensity value they are almost identical in value to the white background. Considering when the RGB channels of a pixel are not equal in value may be able to improve masking of such features. Another possible approach to solving this is to threshold and mask on subsets of the bounding box, as to ensure that very dark trunk pixels do not affect the thresholding of lighter regions.

Our ultimate goal is to create a generalized process that works on classes of specimen images. For the BGNN project we are beginning with fish images, but we are designing the metadata generation system such that it can eventually operate on other species if trained that way. To accomplish this, a much larger training dataset full of more diverse images will be required. The first step towards greater generality will be to operate on other fish collections besides INHS, which is something our program has already show itself capable of doing during initial testing. Another requirement will be to generalize the ruler reading process beyond the INHS specific reading of digits on the ruler, which will likely involve an automated method of reading ruler ticks instead of digits.
% Todo: Are these worth mentioning?
%Adding things like seahorses and eels.
%Fitting a ``spine'' to specimen.
%\section{Conclusion}
%\textbf{Todo}

>>>>

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\linewidth]{images/none1_crop} \\
  \vspace{5mm}
  \includegraphics[width=0.9\linewidth]{images/double1_crop}
  \caption{A fish that was not detected (top) and a fish that was detected twice (bottom).}
  \label{fig:fish_detect_wrong}
\end{figure}

@@

\section{Discussion}
Overall our results show a proof of concept and offer a path forward for using object detection technology, enhanced by image informatics techniques, to improve and enrich metadata that enables advanced specimen image analysis and investigations of scientific research questions on an unprecedented
scale. Our investigation has thus far focused on fish as the specimen of study. Fish are vertebrate animals (phylum Chordata), with over \(34,000\) known unique species~\cite{fishbase},
with many more likely undiscovered. Species names are merely labels, and the discovery of species variation depends on both genotype and phenotype information. The ability to computationally analyze thousands of images of a single fish species, from different habitats and time periods, can lead to new discoveries that are impossible to pursue with manual methods.
Digital library researchers have been concerned with computationally extracting image features, using content based image retrieval methods. The work by Toress~\cite{torres2006digital}, while over 15 years old, demonstrates the challenges and opportunities to automatically generating useful metadata.
Efforts to integrate such automatic metadata generation methods into digital library workflows and architectures still seem limited.
%content based information retrieval (CBIR) for digital library images\cite{ivanova2011content}
%Link above is from the EU counterpart to JCDL, and trying to weave in, just focuses on color semantics, if we want to get another citation in there.
This is likely due to the diversity of image shapes, sizes and the inconsistent configurations of specimens, labels, rulers, etc.\ within them. Object detection as explored in our research, working with an established architecture, is applicable to the larger world of biodiversity, well beyond fish, to include other fauna and flora, art and artefacts, and other digitized objects made accessible for scientific and scholarly research. Following object
detection, one can apply pixel analysis and informatics methods to compute
many more higher order properties from the initial segmentations.
%can we say something here.

Digital libraries serve to collect, provide access to, and archive rich collections of a wide array of materials. Many digital libraries interconnect with open repositories, supporting FAIR (Findable, Accessible, Interoperable, and Reuseable)~\cite{wilkinson2016fair}. In discussing the future of digital library services, Fox \cite{fox2021building} underscores the need to prepare for and test ML applications. The work presented here, within the context of the BGNN project, demonstrates a clear need for improved metadata associated with specimen image collections. Much effort, time and money have been put into photographing and digitizing physical specimens, but without detailed and complete metadata properties the utility of these repositories for advanced computational analysis and ML is limited. Since it is prohibitively time intensive to generate all the pertinent properties by hand, automated techniques are essential to generating the missing metadata at scale.

\section{Conclusion}
In this paper we presented an automatic metadata generation approach. 
Using ML and image informatics algorithms, it is able to locate, mask and analyze specimens (currently limited to fish) in collection images with a high degree of accuracy. It produces 6 of the 22 core BGNN metadata properties~\cite{leipzig2021biodiversity}, as well as image contrast, bounding boxes, scale and length information. Testing this approach on \(7,244\) images from the INHS dataset~\cite{INHS}, we see that the vast majority of the resulting metadata is correct within a tolerance of a few percentage points, and in some cases contains fewer mistakes than the manually generated validation data. Through further refinement and generalization beyond only INHS images, we aim to create a tool that can be distributed to specimen image collection curators to correct the metadata sparsity that precipitated this work.
\subsection{Future Work}
The most pressing next step is to refine the pixel analysis thresholding process so that the entirety of even light colored fish are marked as foreground in the mask. A deficiency of the current process is that it only operates on single channel intensity. Some of the lightest tails appear yellow in hue to the human eye and easily distinguishable, but when compressed to a single intensity value they are almost identical in value to the white background. Considering when the RGB channels of a pixel are not equal in value may improve masking of such features. Another possible approach to solving this
problem is to threshold and mask on subsets of the bounding box, as to ensure that very dark trunk pixels do not affect the thresholding of lighter regions.

Our ultimate goal is to create a generalized process that works on classes of specimen images. For the BGNN project we are beginning with fish images, but we are designing the metadata generation system so that it can eventually operate on other species if appropriately  trained. To accomplish this, a much larger training dataset consisting of more diverse images will be required. The first step towards greater generality will be to operate on other fish collections besides INHS, which is something our program has already
shown itself capable of doing during initial testing. Another requirement will be to generalize the ruler reading process beyond the INHS-specific reading of digits on the ruler, which will likely involve an automated method of reading ruler ticks instead of digits.
% Todo: Are these worth mentioning?
%Adding things like seahorses and eels.
%Fitting a ``spine'' to specimen.
%\section{Conclusion}
%\textbf{Todo}
